---
title: "NYPD_Data"
author: "T. Grey"
date: "2024-10-15"
output:
  pdf_document: default
  html_document: default
---

## Introduction

In this analysis, we are exploring the dataset provided by NYC Open Data from the following source:

[NYC Open Data - Link to the dataset]

<https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD>

### Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE}

library(tidyverse)
library(lubridate)
library(janitor)
library(corrplot)

url <- "https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD"
nyc_data <- read_csv(url)

nyc_data <- janitor::clean_names(nyc_data)

head(nyc_data)
summary(nyc_data)
str(nyc_data)
colnames(nyc_data)

colSums(is.na(nyc_data))

nyc_data <- nyc_data %>%
  mutate(occur_date = mdy(occur_date))

nyc_data <- nyc_data %>%
  mutate(
    hour_of_day = hour(occur_date),                       
    day_of_week = wday(occur_date, label = TRUE),         
    boro_num = as.numeric(factor(boro))                   
  )

# Plot number of incidents over time
nyc_data_summary <- nyc_data %>%
  group_by(occur_date) %>%
  summarise(incident_count = n(), .groups = 'drop')

ggplot(nyc_data_summary, aes(x = occur_date, y = incident_count)) +
  geom_line(color = "darkgreen") +
  labs(title = "Number of Incidents Over Time", 
       x = "Date", 
       y = "Number of Incidents") +
  theme_minimal()

# Histogram of incident counts
numeric_cols <- nyc_data_summary %>% select_if(is.numeric)
if (nrow(numeric_cols) > 0) {
  ggplot(nyc_data_summary, aes(x = incident_count)) + 
    geom_histogram(binwidth = 10, fill = "steelblue", color = "white") +
    labs(title = "Distribution of Incidents", x = "Incident Count", y = "Frequency")
} else {
  print("No numeric columns found for histogram.")
}

boroughs_to_compare <- c("MANHATTAN", "BROOKLYN")

filtered_data <- nyc_data %>%
  filter(boro %in% boroughs_to_compare)

incident_counts <- filtered_data %>%
  group_by(boro) %>%
  summarise(incident_count = n(), .groups = 'drop')  

# Visualize the frequency of incidents by borough
ggplot(incident_counts, aes(x = reorder(boro, -incident_count), y = incident_count, fill = boro)) +
  geom_bar(stat = "identity") +
  labs(title = "Frequency of Incidents by Borough", 
       x = "Borough", 
       y = "Incident Count") +
  theme_minimal() +
  theme(legend.position = "none")

# Correlation analysis
filtered_incident_counts <- filtered_data %>%
  group_by(boro) %>%
  summarise(incident_count = n(), 
            hour_of_day = hour(occur_date),
            boro_num = as.numeric(factor(boro)),
            .groups = 'drop')

correlation_data <- filtered_incident_counts %>%
  select(incident_count, hour_of_day, boro_num)

if (nrow(correlation_data) > 0) {
  cor_matrix <- cor(correlation_data, use = "complete.obs")
  
  # Visualize the correlation matrix
  corrplot(cor_matrix, method = "circle", type = "lower", 
           title = "Correlation Matrix of Incidents Data")
} else {
  print("No numeric columns found for correlation analysis.")
}
